Assessing primitives performance\\ on multi-stage execution

### 2018-05-18 paper.tex
update bench explanation. the LoC was not different from before

\subsection{Results} 
\cba{Please precise results. x\% for this size showing the problem. y\% for other size so non-existent problem. }

We performed the benchmarks upon five different string sizes: 0, 1, 5, 100 and 1000 characters, using two different compilers: GCC on one hand and LLVM on the other hand. Figure~\ref{fig:bench} shows the results in terms of relative speed-up towards the baseline implementation, respectively for GCC and LLVM.

We can draw the following patterns from these results:
\begin{itemize}
\item The execution time of the Plugin implementations is either very close or greater than \sd{within which ranges? 10\% 200\%} the baseline's for the small strings.
\item The Slang+JIT implementation is more efficient than the baseline in every case. Nevertheless, the performance gap between this version and the PureSlang one \sk{all the Slang versions, in the end} tends to decrease as the string size increases. 
\item \sk{if we keep these results, then it actually leads to huge differences...} The compiler used (either GCC or LLVM) does not lead to significant differences in terms of performance, except for the 100+-characters strings. 
\end{itemize}

@@

\paragraph{Stack switch overhead.}
As mentioned in Section~\ref{sec:VMexec}, all implementations written in Slang come with an overhead due to the switch between the C and Smalltalk stacks. This overhead is clearly visible in the bench results in Figure~\ref{fig:bench} for small strings (size 1 or less): the Plugin implementation is around 30 to 60\% slower than the baseline's, and the PureSlang performance is close to the baseline's.
As the string size goes larger, the overhead is less significant because it is absorbed by other performance gains.
Regarding this criteria, the Slang+JIT version is always the fastest as it is not impacted by the switch overhead.

\paragraph{Optimization potential on large strings comparison.}
The plugins come with an overhead due the use of an interface. Besides, cross-file inlining is not performed for plugins, like said in Section~\ref{sec:implem_cmp}. These are some of the factors explaining the performance gap between these versions and the Pure Slang and Slang+JIT ones.
When it comes to comparing large strings (100 characters and more), the execution time is mainly spent in the comparison loop (Figure~\ref{fig:Asm}). However, the switch overhead is still noticeable in the PureSlang implementation when strings of 100 characters are compared: PureSlang is 15.8 times faster than Baseline while Slang+JIT is 16.8 times faster than Baseline. This is not the case anymore on 1000-characters strings.
This last observation only applies to the GCC results: the code generated by the LLVM compiler induces a longer execution time, as explained in the next paragraph.

One could notice a rather striking performance difference between the 2 compilers: when compiling with GCC, the performance between the PureSlang and Slang+JIT implementations is really close for long strings (100 characters and longer), with respectively 24,42 and 24,48 relative speed-up factors for a 1000-characters string. Yet performance greatly differs when LLVM is used, with respectively 13,44 and 26,40 relative speed-up factors.

@@

\todo{add conclusion SK}

>>>>

\subsection{Results} 
We performed the benchmarks upon five different string sizes: 0, 1, 5, 100 and 1000 characters, using two different compilers: GCC on one hand and LLVM on the other hand. Figure~\ref{fig:bench} shows the results in terms of relative speed-up towards the baseline implementation, respectively for GCC and LLVM.

We can draw the following patterns from these results:
\begin{itemize}
\item The execution time of the Plugin implementations is greater than the baseline's for empty strings (0,58 times slower) or 1-character strings (0,23 times slower).
\item The Slang+JIT implementation is more efficient than the baseline in every case. Nevertheless, the performance gap between this version and the Plugin and Slang ones tends to decrease as the string size increases, especially for GCC. For 1000-characters strings, the Slang+JIT version goes 0,25 times faster than the Plugin and the Slang versions.
\item The compiler used (either GCC or LLVM) does lead to significant differences in terms of performance for the Slang implementation. 
\end{itemize}

@@

\paragraph{Stack switch overhead.}
As mentioned in Section~\ref{sec:VMexec}, all implementations written in Slang come with an overhead due to the switch between the C and Smalltalk stacks. This overhead is clearly visible in the bench results in Figure~\ref{fig:bench} for small strings (size 1 or less): the Plugin implementation is between 57,8 (gcc) to 61,4\% (llvm) slower than the baseline's for empty strings.
As the string size goes larger, the overhead is less significant because it is absorbed by other performance gains.
Regarding this criteria, the Slang+JIT version is always the fastest as it is not impacted by the switch overhead.

\paragraph{Optimization potential on large strings comparison.}
The Plugin come with an overhead due the use of an interface. Besides, cross-file inlining is not performed for plugins, like said in Section~\ref{sec:implem_cmp}. These are some of the factors explaining the performance gap between these versions and the Slang and Slang+JIT ones.
When it comes to comparing large strings (100 characters and more), the execution time is mainly spent in the comparison loop (Figure~\ref{fig:Asm}). However, the switch overhead is still noticeable in the Slang implementation when strings of 100 characters are compared: Slang is 17.7 times faster than Baseline while Slang+JIT is 18,6 times faster than Baseline. 
This last observation only applies to the GCC results: the code generated by the LLVM compiler induces a longer execution time, as explained in the next paragraph.

\paragraph{Control over generated code.} One could notice a rather striking performance difference between the 2 compilers: when compiling with GCC, the performance between the Slang and Slang+JIT implementations is really close for long strings (100 characters and longer), with respectively 25,42 and 26,76 relative speed-up factors for a 1000-characters string. Yet performance greatly differs when LLVM is used, with respectively 10,67 and 25,4 relative speed-up factors.

@@

One can also notice the gap between the Plugin and Slang implementations in LLVM, which does not appear in GCC. \sk{one last comment: WHY?!}

In this section, we compared the performance of the different implementations against a baseline (\ie{pure Smalltalk code}) implementation. The results showed the Slang+JIT implementation was always the fastest. Otherwise, most of the performance gaps between the primitive versions could be explained by 3 criteria: stack swithing, optimization for large strings, and the control over generated code.
